### 2019-11-18 paper/paper.tex
Finished section on Fig 1

With the addition of the competition term, $\chi$, it is necessary to
introduce a normalisation step, in order to prevent one TC type from
out-competing every other TC type. Normalisation conforms to the idea that for
a fixed number of afferent thalamocortical axons, the overall number of
branches must have some upper limit; given the presumably finite resources
available in each axon. We divisively normalize the branching density,
dividing $a_i$ by the mean of $a_i$ per unit area; formalised as:
%
\begin{equation} \label{eq:norm}
a_i(\mathbf{x}, t) = A \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega a'_i(\mathbf{x}, t) d\sigma}
\end{equation}
%
where the prime symbols indicate the use of the intermediate value of $a_i$
after it has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:branch_comp}
on the basis of the system state at time $t-{\delta}t$. $A$ is the area of the
domain. This normalises $a_i(\mathbf{x}, t)$ to 1, integrated over the domain. To
account for the possibility that there might be more axons afferent for TC
type $1$ cf.~TC type $2$, we normalise to the integral of $a_i$ at time 0.
%
\begin{equation} \label{eq:norm_time0}
\begin{split}
a_i(\mathbf{x}, t) & =
\frac{\int_\Omega  a'_i(\mathbf{x}, 0) d\sigma}{A} \; A \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
  a'_i(\mathbf{x}, t) d\sigma} \\
%
& = \int_\Omega  a'_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
  a'_i(\mathbf{x}, t) d\sigma}
\end{split}
\end{equation}
%
In the simulation, in which space is represented as $n_e$ discrete
hexagonal tiles, the equivalent expression is:
%
\begin{equation} \label{eq:norm_discrete}
{a_{i,p}}(t) = \textstyle \sum_j^{n_e} a_{i,j}(0) \; \frac {a'_{i,p}(t)} {\textstyle \sum_j^{n_e} a'_{i,j}(t) }
\end{equation}
%
where $a_{i,j}(t)$ is the value of the variable $a_i$ in the $j^{\mathrm{th}}$
hexagonal element at time t.

We first investigated the effect of the competition by examining a system of
25 thalamocortical axon types simulated on a square domain, as in
Fig.~\ref{fig:square}. The hex to hex distance was 0.02 mm and the
square's side length was 1.16 mm. The axons interacted with two orthogonal
signalling gradients, one of which increased from the top edge to the bottom
edge, the other of which had its gradient increasing from the right edge to
the left edge [Fig.~\ref{fig:square}(f)]. For each thalamocortical type, $i$,
two interaction parameters were set; $\gamma_{i,1}$ was the interaction with
guidance molecule 1 and $\gamma_{i,2}$ was the interaction with molecule
2. The branching of TC types with a negative interaction with molecule 2
($\gamma_{i,2}<0$) will be pushed to the left; those with $\gamma_{i,2}>0$ are
pushed to the right. TC types that interact negatively with molecule 1
($\gamma_{i,1}<0$) are pushed up; those with $\gamma_{i,1}>0$ are pushed down.

Fig.~\ref{fig:square}(a) shows the connection densities after the first 100
simulation time steps (t=0.005 s). Each colour uniquely identifies the TC
type, $i$ [the full mapping between index $i$ and the colour is shown in
  Fig.~\ref{fig:square}(d)]. The left-hand map colours each hex according to
which TC type has the highest connection density. At this early stage of the
simulation, the colouring is somewhat random because the branching density is
initialised stochastically ($a_{i,j}(0) = 0.2 + U([0,0.2])$), which results in
a randomised pattern for ${c_i}^{max}$. The right-hand map of
Fig.~\ref{fig:square}(a) colours those hexes which have $c_i \geq 0.5$ and an
adjoining hex for which $c_i < 0.5$ according to the mapping shown in
Fig.~\ref{fig:square}(d)]. Due to the initial stochasticity, there are many
  such crossings of the threshold. The predominance of blue (for $i<10$)
  results from the order in which the hexes have been rendered.
Fig.~\ref{fig:square}(b) shows the simulation at t=0.14. \cmnt{CONTINUE FROM HERE.}

% Later
The full set of these interaction parameters are shown in
Fig.~\ref{fig:square}(e). The values of $\gamma$ were chosen with the
intention of forming a grid; the interactions for $i =$ 1, 5, 21 \& 25 (the
corners) were strongest and the interaction parameters for $i=$13 (the centre)
were zero. Fig.~\ref{fig:square}(d) shows the mapping between colour and TC
type $i$ and can be used, in conjunction with Fig.~\ref{fig:square}(e), to see
that $\gamma_{1,1}=-2$, $\gamma_{1,2}=-2$, $\gamma_{3,1}=0$,
$\gamma_{3,2}=-2$, and so on.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{./Fig_25N2M_square/Fig_25N2M_square.png}
\end{center}
%\vspace{-0em}
\caption{\textbf{(a)--(e)} Connection densities of 25 thalamocortical axons
  simulated on a square domain. \textbf{(a)} Connection density at simulation
  time t=0. The left image shows a map of the identity of the thalamocortical
  axon type with the highest absolute connection density. Hexagonal elements
  coloured the darkest blue have the highest connection density for TC type
  $i=1$; those coloured the darkest red have maximal connection density in TC
  type $i=25$. The right hand image of each pair shows the hexes for which the
  connection density passes 0.5 for each TC type, giving a contour map. At
  t=0, connection densities are randomly distributed. \textbf{(b)} t=0.14. The
  grid-like organisation of the TC connections is beginning to emerge.
  \textbf{(c)} t=0.2. The 5x5 grid is now clear in the left map. Branching
  competition and differential interaction with the two guidance gradients is
  causing the branching, and therefore the connection densities, to localise in
  25 different locations. \textbf{(d)} t=0.64. Connection densities for each
  TC type have increased sufficiently for the contours to form. In (d), the
  mapping between colour and TC type $i$ is indicated. \textbf{(e)}
  t=1.41. Contours are well formed. In (e), the $\gamma$ values for each TC
  type, $i$, are shown. $\gamma_{i,1}$ gives the interaction with molecule `1'
  [red map in (f)]; $\gamma_{i,2}$ is the interaction with molecule
  `2'. \textbf{(f)} The two guidance gradients (`1' and `2') for this
  simulation.}
\label{fig:square}
\end{figure*}

>>>>

With the addition of the competition term, $\chi$, it is necessary to
introduce a normalisation step, in order to prevent one TC type from
out-competing every other TC type. Normalisation represents the idea that for
a fixed number of afferent thalamocortical axons, the overall number of
branches must have some upper limit; given the finite resources available in
each axon for building new processes. We divisively normalize the branching
density, dividing it at each computational step by the mean branching density
per unit area; formalised as:
%
\begin{equation} \label{eq:norm}
a_i(\mathbf{x}, t) = A \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega a'_i(\mathbf{x}, t) d\sigma}
\end{equation}
%
where the prime symbols indicate the use of the intermediate value of $a_i$
after it has been computed according to Eqs.~\ref{eq:dc}--\ref{eq:branch_comp}
on the basis of the system state at time $t-{\delta}t$. $A$ is the area of the
domain. This normalises $a_i(\mathbf{x}, t)$ to 1, integrated over the domain. To
account for the possibility that there might be more axons afferent for TC
type $1$ cf.~TC type $2$, we normalise to the integral of $a_i$ at time 0, so
that Eq.~\ref{eq:norm} becomes:
%
\begin{equation} \label{eq:norm_time0}
\begin{split}
a_i(\mathbf{x}, t) & =
\frac{\int_\Omega  a_i(\mathbf{x}, 0) d\sigma}{A} \; A \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
  a'_i(\mathbf{x}, t) d\sigma} \\
%
& = \int_\Omega  a_i(\mathbf{x}, 0) d\sigma \; \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
  a'_i(\mathbf{x}, t) d\sigma}
%\\
%& = a_0 \frac {a'_i(\mathbf{x}, t)} {\int_\Omega
%  a'_i(\mathbf{x}, t) d\sigma}
\end{split}
\end{equation}
%
In the simulation, in which space is represented as $n_e$ discrete
hexagonal tiles, the equivalent expression is:
%
\begin{equation} \label{eq:norm_discrete}
  {a_{i,p}}(t) = \textstyle \sum_j^{n_e} a_{i,j}(0) \; \frac {a'_{i,p}(t)}
  {\textstyle \sum_j^{n_e} a'_{i,j}(t) }
\end{equation}
%
where $a_{i,j}(t)$ is the value of the variable $a_i$ in the $j^{\mathrm{th}}$
hexagonal element at time t.

We first investigated the effect of the competition by examining a system of
25 thalamocortical axon types simulated on a square domain, as in
Fig.~\ref{fig:sq}, to determine if a complex pattern could emerge. The hex
to hex distance was 0.02 mm and the square's side length was 1.16 mm (3919 hex
elements). The axons interacted with two orthogonal signalling gradients, one
of which increased from the top edge to the bottom edge, the other of which
had its expression increasing from the right edge to the left edge
[Fig.~\ref{fig:sq}(f)]. For each thalamocortical type, $i$, two
interaction parameters were set; $\gamma_{i,1}$ was the interaction with
guidance molecule 1 and $\gamma_{i,2}$ was the interaction with molecule
2. The branching of TC types with a negative interaction with molecule 2
($\gamma_{i,2}<0$) were pushed to the left; those with $\gamma_{i,2}>0$ were
pushed to the right. TC types that interacted negatively with molecule 1
($\gamma_{i,1}<0$) were pushed up; those with $\gamma_{i,1}>0$ were pushed
down.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{./Fig_25N2M_square/Fig_25N2M_square.png}
\end{center}
%\vspace{-0em}
\caption{\textbf{(a)--(e)} Connection densities of 25 thalamocortical axons
  simulated on a square domain. \textbf{(a)} Connection density at simulation
  time t=0.005. The left image shows a map of the identity of the thalamocortical
  axon type with the highest absolute connection density. Hexagonal elements
  coloured the darkest blue have the highest connection density for TC type
  $i=1$; those coloured the darkest red have maximal connection density in TC
  type $i=25$. The right hand image of each pair shows the hexes for which the
  connection density passes 0.5 for each TC type, giving a contour map. At
  t=0, connection densities are randomly distributed. \textbf{(b)} t=0.18. The
  grid-like organisation of the TC connections is beginning to emerge.
  \textbf{(c)} t=0.25. The 5x5 grid is now clear in the left map. Branching
  competition and differential interaction with the two guidance gradients is
  causing the branching, and therefore the connection densities, to localise
  in 25 different locations. \textbf{(d)} t=0.81. Connection densities for
  each TC type have increased sufficiently for the contours to form. In (d),
  the mapping between colour and TC type $i$ is indicated. \textbf{(e)}
  t=1.77. Contours are well formed. In (e), the $\gamma$ values for each TC
  type, $i$, are shown. $\gamma_{i,1}$ gives the interaction with molecule `1'
  [red map in (f)]; $\gamma_{i,2}$ is the interaction with molecule
  `2'. \textbf{(f)} Expression maps for the two guidance molecules (`1' and
  `2') for this simulation. Colour saturation indicates expression intensity,
  which for each molecule varies linearly in one direction.}
\label{fig:sq}
\end{figure*}

Fig.~\ref{fig:sq}(a) shows the connection densities after the first 100
simulation time steps (t=0.005 s). Each colour uniquely identifies the TC
type, $i$ [the full mapping between index $i$ and the colour is shown in
  Fig.~\ref{fig:sq}(d)]. The left-hand map colours each hex according to
which TC type has the highest connection density. At this early stage of the
simulation, the colouring is somewhat random because the branching density was
initialised stochastically ($a_{i,j}(0) = 0.2 + U([0,0.2])$), which resulted
in a randomised pattern for ${c_i}^{max}$. The right-hand map of
Fig.~\ref{fig:sq}(a) colours those hexes which have $c_i \geq 0.5$ and an
adjoining hex for which $c_i < 0.5$ according to the mapping shown in
Fig.~\ref{fig:sq}(d)]. Due to the initial stochasticity, there are many
  such crossings of the threshold. The predominance of blue (for $i<10$)
  results from the order in which the hexes were rendered, with blue hexes
  being coloured last.  Fig.~\ref{fig:sq}(b) shows the simulation at
  t=0.14.
%
%
Fig.~\ref{fig:sq}(b) shows the connection density after 0.18 s of simulation
time (3600 steps of 5x10$^{-5}$ s). The initial randomness in the pattern has
disappearing, to be replaced by coalesced regions in which each TC type $i$ is
maximal.  There is a clear blue/red division from left to right. The threshold
plot indicates that the guidance molecules have pushed the branching towards
the corners of the square for the 4 TC types with the strongest interactions
with the guidance molecules; these exceed the threshold of 0.5. For the other
21 TC types, the connection densities are not yet sufficiently localised to
cross the threshold.
%
0.08 s of simulation time later, in Fig.~\ref{fig:sq}(c), the 5x5 grid is
clearly visible in the maximal value map, though the localisation of each
connection density map is still not sufficient for the contour map to emerge.
%
The contour map emerges by 0.81 s and is well formed by 1.77 s
[Figs.~\ref{fig:sq}(d) \& (e)]. Fig.~\ref{fig:sq}(e) shows the interaction
parameters for each TC type ($\gamma$). The values of $\gamma$ were chosen
with the intention of forming a grid; the interactions for $i =$ 1, 5, 21 \&
25 (the corners) were strongest and the interaction parameters for $i=$13 (the
centre) were zero. Fig.~\ref{fig:sq}(d) shows the mapping between colour and
TC type $i$ and can be used, in conjunction with Fig.~\ref{fig:sq}(e), to see
that $\gamma_{1,1}=-2$, $\gamma_{1,2}=-2$, $\gamma_{3,1}=0$,
$\gamma_{3,2}=-2$, and so on.

The simulation shown in Fig.~\ref{fig:sq} demonstrates that complex structure
can form in a system with just two guidance gradients \cmnt{(provided that
  interactions with each gradient occur in both senses; positive and negative)}.
